Topological Equivariant Artist Model for Visualization Library Architecture

### 2021-11-22 paper/main.tex
sections

\begin{definition} The \textcolor{fiber}{fiber space} \dfiberc\ is a topological space such that for every $\dbasepoint \in \dbase$, the fiber over that point is isomorphic to the preimage of that point $\dfiber_{\dbasepoint} \cong \pi^{-1}(\{\dbasepoint\})$. All fibers $\dfiber_{\dbasepoint}$ are homeomorphic to each other \cite{spanier1989algebraic, hatcherAlgebraicTopology2002}
\end{definition}
\note{intuition about fiber}

@@

\begin{equation}
  \begin{tikzcd}
      \dfiberc \arrow[r, hook] & \dtotalc \arrow[d, "\pi"'] \\
                        & \dbasec \arrow[u, "\dsectionc"', bend right]
  \end{tikzcd}
\end{equation}

How these spaces fit together is illustrated in \autoref{fig:related-work:fiber-bundle}, wherein values lives in the fiber $\dfiber\subseteq\dtotal$. In this example, the fiber \dfiber\ is the cartesian product of two sets $F_{0}\times F_{1}$ where each fiber $F_{i}$ is ....

\subsubsection{Sheaves}
The set of all sections sections $\{\dsection^{i}: \openset_{j}\rightarrow \dtotal\}|_{i\in I}$ over an arbitrary open set $U_{j}$ is denoted $\Gamma(\openset_{\dbasepoint}, \dtotal)$. The set of all sections over the total space \dbase\ is denoted $\Gamma(\dbase, \dtotal)$.

The map from the open set to the set of sections is called a presheaf $\sheaf: U_j \rightarrow \Gamma(\openset_j, \dtotal)$ \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}. The presheaf preserves inclusion maps between the open sets open sets $\iota$ and the inclusion map $\iota^*$ between the sets of sections.

@@

\subsection{Categorical Artist}
\label{sec:artist:category}

@@

We also introduce a category of open sets called $\mathcal{\gbase}$ that encapsulates the topology of the graphic display space. The category $\mathcal{\gtotal}$ has  inclusion morphisms $\iota: {\openset_{i}}^{\prime} \hookrightarrow {\openset_{j}}^{\prime}$, including into the total set \gtotal. We introduce a surject map $\vindex: \gtotal \rightarrow \dtotal$ from the the graphic base space to the data base space. Given the morphism $\iota$
\begin{equation}
  \label{eq:artist:vindex}
  \begin{tikzcd}
    & \opensetc^{\prime}_i \arrow[d, "\iota"] \\
  \dbasec \arrow[ru, "\vindexc^*", dashed] & \gbasec \arrow[l, "\vindexc"]   
\end{tikzcd}
\end{equation}
we propose that the map \vindex\ can be pulled back to a map between open sets in \dbase and open sets in \gbase. This map is the functor
\begin{equation}
  \vindex^{*}:\mathcal{\dtotal} \rightarrow \mathcal{\gtotal}
\end{equation}
that maps from the data base space to the graphic base space. The map $\vindex^{*}$ is a lookup map between where a point is located in the data and where it is located on screen.

@@

The category $\mathcal{H}$ has as objects the sets of functions that generate every arbitrary graphic over the open sets in $\mathcal{\gtotal}$. While $\mathcal{H}$ is the domain of the artist \vartist, the range of \vartist\ is the subset of graphics $\sheaf_{\vartist} \subset \sheaf_{\mathcal{\dtotal}}$ such that structure imposed by $\dfunc_{\dtotal}$ commutes 
\begin{equation}
  \label{eq:artist:equivariance}
  \begin{tikzcd}
    \sheafc_{\dtotalc} \arrow[r, "\vartistc"] \arrow[d, "\dfuncc_{\dtotalc}"'] & \sheafc_{\vartistc} \arrow[d, "\dfuncc_{\vartistc}"] \\
    \sheafc_{\dtotalc} \arrow[r, "\vartistc"] & \sheafc_{\vartistc}                                              
  \end{tikzcd}
\end{equation}
The equivariance condition expressed in \autoref{eq:artist:equivariance} is that a graphic generated by  transforming and then visualizing data $\vartist(\dfunc_{\dtotal}(\sheaf_{\mathcal{dtotal}}))$ can equivalently be generated by visualizing the data and then transforming the graphic $\dfunc_{\vartist}(\vartist(\sheaf_{\mathcal{\dtotal}}))$. This is a formal statement of the equivariance illustrated in \autoref{fig:related-work:equivariance}

In \autoref{sec:artist:construction}, we show a method of constructing the artist such that the constraints of \autoref{eq:artist:morphism-of-sheaf} and \autoref{eq:artist:equivariance} can be satisfied, thereby ensuring equivariance and the preservation of continuity. We then use this formulation to guide the development of visualization library components in \autoref{sec:case-study}.

>>>>

\begin{definition} The \textcolor{fiber}{fiber space} \dfiberc\ is a topological space such that for every $\dbasepoint \in \dbase$, the fiber over that point is isomorphic to the preimage of that point $\dfiber_{\dbasepoint} \cong \pi^{-1}(\{\dbasepoint\})$. All fibers $\dfiber_{\dbasepoint}$ are homeomorphic to each other \cite{spanier1989algebraic, hatcherAlgebraicTopology2002}
\end{definition}
\note{intuition about fiber + spivak}

@@

\begin{equation}
  \label{eq:related-work:continuity:fiber-bundle}
  \begin{tikzcd}
      \dfiberc \arrow[r, hook] & \dtotalc \arrow[d, "\pi"'] \\
                        & \dbasec \arrow[u, "\dsectionc"', bend right]
  \end{tikzcd}
\end{equation}

\note{this needs to be reworked into a coherent sentence}
How these spaces fit together is illustrated in \autoref{fig:related-work:fiber-bundle}, wherein values lives in the fiber $\dfiber\subseteq\dtotal$. In this example, the fiber \dfiber\ is the cartesian product of two sets $F_{0}\times F_{1}$ where each fiber $F_{i}$ is ....

As proposed by Butler et al, data can be represented as a map from the base space \dbase\ to the fiber space \dfiber. This map is called a section $\dsection: \dbase \rightarrow \dtotal$ of the fiber bundle. A fiber bundle has many sections, and the set of all sections
\begin{equation}
\label{eq:related-work:continuity:sections}
\{\dsection^{i}: \openset_{j}\rightarrow \dtotal\}|_{i\in I} = \Gamma(\openset_{j}, \dtotal)
\end{equation}
over a subset \openset\ is the set of all data sets with a shared fiber \dfiber\ and continuity \dbase. The set of all sections defined over the total base space is denoted $\Gamma(\dbase, \dtotal)$.

\subsubsection{Sheaves}
The map from the open set to the set of sections is called a presheaf $\sheaf: U_j \rightarrow \Gamma(\openset_j, \dtotal)$ \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}. The presheaf preserves inclusion maps between the open sets open sets $\iota$ and the inclusion map $\iota^*$ between the sets of sections.

@@

\subsection{Natural Transformation of Data to Graphic}
\label{sec:artist:category}

@@

Over these open sets $\openset_i$ are sets of local sections

@@

We introduce a surjective map $\vindex: \gtotal \rightarrow \dtotal$ from the the graphic base space to the data base space. Given the morphism $\iota$
\begin{equation}
  \label{eq:artist:vindex}
  \begin{tikzcd}
    & \opensetc^{\prime}_i \arrow[d, "\iota"] \\
  \dbasec \arrow[ru, "\vindexc^*", dashed] & \gbasec \arrow[l, "\vindexc"]   
\end{tikzcd}
\end{equation}
we propose that the map \vindex\ can be pushed forward to a map between open sets in \dbase and open sets in \gbase. This map is the functor
\begin{equation}
  \vindex_{*}:\mathcal{\dtotal} \rightarrow \mathcal{\gtotal}
\end{equation}
that maps from open sets $\openset_i \subset \dbase$ in data space to the open sets in visual space $\openset^{\prime}_i \subset \gbase$. The graphic space is

\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{pushforward.png}
  \label{artist:pushforward_xi}
  \caption{The pushforward map $\vindex_*$ repeats the 1D fiber over every point $\dbasepoint \in \dbase$ in data space, turning the fiber into a 2D plane of repeating values over the corresponding region $\dbasepoint = \vindex(\gbasepoint)$ in the graphic space \gbase. }
\end{figure}
